---
title: "RShiny Top 20 Drug Analysis Code"
author: "Jonathan Harris"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ProPublica Top 50 Drug Dataset Analysis

##Setup DataFrames for Plots and RShiny
### Setup Program
```{r, message=FALSE, warning=FALSE}

#Clearout environment
rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
gc()

#Load Packages
library(tidyverse)
library(dplyr)
library(ggplot2)
library(scales)
library(RColorBrewer)
library(broom)
library(ggalt)

#Download BBC Theme
if(!require(pacman))install.packages("pacman")
pacman::p_load('dplyr', 'tidyr', 'gapminder',
               'ggplot2',  'ggalt',
               'forcats', 'R.utils', 'png', 
               'grid', 'ggpubr', 'scales',
               'bbplot')
devtools::install_github('bbc/bbplot')

setwd("/Users/jonathanharris/Desktop/NYCDSA/projects/rshiny_proj")

```

### Import Data
```{r data import}

drug_df = read_csv("DataSets/ProPub_TopDrugs/doctor_drug_2016.csv")
head(drug_df)

```

### ID Most Prescribed/Expensive Drugs
```{r top drugs}

#ID top 50 most prescribed/expensive drug and Isolate Drugs with Most Payments
topdrugs = drug_df %>% filter(top50_claims == T &
                                top50_cost == T) %>%
  group_by(drug_name) %>%
  summarise(tot = sum(d4d_payment_amount)) %>%
  arrange(desc(tot)) %>% select(drug_name) %>% head(20)

#Isolate drug_df
drugs_interest = drug_df %>% filter(drug_name %in% unlist(topdrugs))

#Import topdrug descriptions
drugs_describe = read_csv("DataSets/ProPub_TopDrugs/drug_id.csv")

#InnerJoin with drugs_interest
drugs_interest = left_join(drugs_interest, drugs_describe, by = 'drug_name') %>% drop_na %>% mutate(drug_name = str_to_title(drug_name))
rm(drug_df) #remove original dataframe for RAM

head(drugs_interest)

```

### Prep Data for Shiny
```{r rshiny data prep}

#Prepare/export drug df for rshiny app
rshiny_drugs = drugs_interest %>% select(-year ,
                                         -top50_claims,
                                         -top50_cost,
                                         -generalist,
                                         -total_claim_count_other,
                                         -drug_id) %>%
  rename(
    'received_payments' = 'd4d_hasPayment',
    'n_payments' = 'd4d_payment_number',
    'tot_payments' = 'd4d_payment_amount',
    'tot_meal_payments' = 'd4d_payment_amount_meals',
    'tot_nonmeal_payments' = 'd4d_payment_amount_nmeals',
    'tot_promo_payments' = 'd4d_payment_amount_ps',
    'meet_claim_thresh' = 'partD_hasClaim',
    'n_claims_drug' = 'partD_claim_count',
    'tot_claim_drug' = 'partD_claim_cost',
    'n_claims_alldrugs' = 'total_claim_count',
    'tot_claims_alldrugs' = 'total_drug_cost',
  ) %>%
  mutate(payment_type = ifelse(
    tot_payments == 0,
    'NoPayments',
    ifelse(
      tot_meal_payments > 0 &
        tot_nonmeal_payments == 0 & tot_promo_payments == 0,
      'MealOnly',
      ifelse(
        tot_meal_payments == 0 &
          tot_nonmeal_payments > 0 & tot_promo_payments == 0,
        'NonMealOnly',
        ifelse(
          tot_meal_payments == 0 &
            tot_nonmeal_payments == 0 &
            tot_promo_payments > 0,
          'PromoOnly',
          'AllTypes'
        )
      )
    )
  ))

write_delim(rshiny_drugs,
            path = "NYCDSA_doctorpayments_jonharris_rshinyapp/rshiny_drugs.csv",
            delim = ';',
            col_names = T)

#Prepare/export drug description df for rshiny app
rshiny_drug_type = data.frame(cbind(drugs_interest$drug_name, drugs_interest$drug_id)) %>%
  distinct() %>%
  rename('drug_name' = 'X1', 'drug_type' = 'X2') %>%
  mutate(drug_name = str_to_title(drug_name),
         drug_type = str_to_title(drug_type))
write_delim(
  rshiny_drug_type,
  path = "NYCDSA_doctorpayments_jonharris_rshinyapp/rshiny_drugtype.csv",
  delim = ';',
  col_names = T
)

rm(rshiny_drugs) #remove rshiny df for RAM

```

### Add Additional Preparation
```{r additional prep}

drugs_interest = drugs_interest %>% select(-year ,
                                           -top50_claims,
                                           -top50_cost,
                                           -generalist,
                                           -total_claim_count_other) %>%
  rename(
    'received_payments' = 'd4d_hasPayment',
    'n_payments' = 'd4d_payment_number',
    'tot_payments' = 'd4d_payment_amount',
    'tot_meal_payments' = 'd4d_payment_amount_meals',
    'tot_nonmeal_payments' = 'd4d_payment_amount_nmeals',
    'tot_promo_payments' = 'd4d_payment_amount_ps',
    'meet_claim_thresh' = 'partD_hasClaim',
    'n_claims_drug' = 'partD_claim_count',
    'tot_claim_drug' = 'partD_claim_cost',
    'n_claims_alldrugs' = 'total_claim_count',
    'tot_claims_alldrugs' = 'total_drug_cost',
  ) %>%
  mutate(payment_type = ifelse(
    tot_payments == 0,
    'NoPayments',
    ifelse(
      tot_meal_payments > 0 &
        tot_nonmeal_payments == 0 & tot_promo_payments == 0,
      'MealOnly',
      'AllTypes'
    )
  )) %>%
  mutate(payment_amt_group = ifelse(
    tot_payments == 0,
    'nopayment',
    ifelse(
      tot_payments > 0 & tot_payments < 100,
      'bet_1_100',
      ifelse(
        tot_payments >= 100 & tot_payments < 500,
        'bet_100_500',
        ifelse(
          tot_payments >= 500 & tot_payments < 10000,
          'bet_500_1k',
          ifelse(
            tot_payments >= 1000 &
              tot_payments < 5000,
            'bet_1k_5k',
            'great_5k'
          )
        )
      )
    )
  )) %>% mutate(drug_name_long = str_to_title(paste(
    paste(drug_name, drug_id, sep = ' ('), ')', sep = ''
  ))) 

```

## Create Plots
### Plot: Number of Physicians
```{r ndoc_plot}

#var = n docs per drug
data = drugs_interest %>% group_by(drug_name) %>%
  select(drug_name, npi)  %>%
  summarise(var = n())

ndoc_plot =
  ggplot(data = data, aes(
    x = reorder(drug_name, -var),
    y = var,
    fill = drug_name
  )) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = ifelse(data$drug_name == 'Oxycontin', '#FAAB18' , '#1380A1')) +
  bbc_style() +
  labs(title = "Number of Prescribing Clinicians, By Drug",
       subtitle = 'Large variation in number of prescribers',
       caption = "Data source: 2016 Top 50 Drugs, propublica.org") +
  scale_y_continuous(labels = scales::comma) +
  geom_hline(yintercept = 0,
             size = 1,
             colour = "#333333") + coord_flip() +
  geom_label(
    aes(label = format(
      round(var, 0), big.mark = ",", scientific = FALSE
    )),
    hjust = 1,
    vjust = 0.5,
    colour = "white",
    fill = NA,
    label.size = NA,
    family = "Helvetica",
    size = 3
  ) +
  theme(
    panel.grid.major.x = element_line(color = "#cbcbcb"),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(1, 1, 1, 1, "cm"),
    legend.position = 'none',
    plot.title = element_text(size = 24),
    plot.subtitle = element_text(size = 16, face = "italic"),
    plot.caption = element_text(size = 6, hjust = 1)
  )

#Save Plot
ggsave(
  'plots/topdrugs_ndocs.png',
  plot = ndoc_plot,
  units = "in",
  height = 7,
  width = 12,
  dpi = 300
)

```

### Plot: Average Number of Claims Per Physician by Drug
```{r avgclaims_plot}

#Var = mean claims per drug
data = drugs_interest %>% group_by(drug_name)  %>%
  summarise(var = mean(n_claims_drug))

avgclaims_plot =  ggplot(data = data, aes(
  x = reorder(drug_name, -var),
  y = var,
  fill = drug_name
)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = ifelse(data$drug_name == 'Oxycontin', '#FAAB18' , '#1380A1')) +
  bbc_style() +
  labs(title = "Average Number of Claims Per Physician, By Drug",
       subtitle = "Maintenance drugs among most prescribed",
       caption = "Data source: 2016 Top 50 Drugs, propublica.org") +
  scale_y_continuous(labels = scales::comma) +
  geom_hline(yintercept = 0,
             size = 1,
             colour = "#333333") + coord_flip() +
  geom_label(
    aes(label = format(
      round(var, 0), big.mark = ",", scientific = FALSE
    )),
    hjust = 1,
    vjust = 0.5,
    colour = "white",
    fill = NA,
    label.size = NA,
    family = "Helvetica",
    size = 3
  ) +
  theme(
    panel.grid.major.x = element_line(color = "#cbcbcb"),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(1, 1, 1, 1, "cm"),
    legend.position = 'none',
    plot.title = element_text(size = 24),
    plot.subtitle = element_text(size = 16, face = "italic"),
    plot.caption = element_text(size = 6, hjust = 1)
  )

ggsave(
  'plots/topdrugs_avgclaims.png',
  plot = avgclaims_plot,
  units = "in",
  height = 7,
  width = 12,
  dpi = 300
)

```

### Plot: Percent of Physicians on Payroll by Drug
```{r percpayroll_plot}

#Var = % of doctors paid by drug companies
data = drugs_interest %>% group_by(drug_name)  %>%
  summarise(var = sum(received_payments) / n())

percpayroll_plot =
  ggplot(data = data, aes(
    x = reorder(drug_name, -var),
    y = var,
    fill = drug_name
  )) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = ifelse(data$drug_name == 'Oxycontin', '#FAAB18' , '#1380A1')) +
  bbc_style() +
  labs(title = "Percentage of Clinicians Receiving Payments",
       subtitle = "Diabetes market hypercompetitive",
       caption = "Data source: 2016 Top 50 Drugs, propublica.org") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2)) +
  geom_hline(yintercept = 0,
             size = 1,
             colour = "#333333") + coord_flip() +
  geom_label(
    aes(label = scales::percent(round(var, 2))),
    hjust = 1,
    vjust = 0.5,
    colour = "white",
    fill = NA,
    label.size = NA,
    family = "Helvetica",
    size = 3
  ) +
  theme(
    panel.grid.major.x = element_line(color = "#cbcbcb"),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(1, 1, 1, 1, "cm"),
    legend.position = 'none',
    plot.title = element_text(size = 24),
    plot.subtitle = element_text(size = 16, face = "italic"),
    plot.caption = element_text(size = 6, hjust = 1)
  )

ggsave(
  'plots/topdrugs_percpayroll.png',
  plot = percpayroll_plot,
  units = "in",
  height = 7,
  width = 12,
  dpi = 300
)

```

### Plot: Average Total Physician Payments by Drug
```{r avgpayment_plot}

#Var = average payment per physician
data = drugs_interest %>% group_by(drug_name)  %>%
  filter(received_payments == T) %>%
  summarise(var = mean(tot_payments))

avgpayment_plot = ggplot(data = data, aes(
  x = reorder(drug_name, -var),
  y = var,
  fill = drug_name
)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = ifelse(data$drug_name == 'Oxycontin', '#FAAB18' , '#1380A1')) +
  bbc_style() +
  labs(title = "Average Clinician Payment, By Drug",
       subtitle = "Large discrepancy suggests different payment types",
       caption = "Data source: 2016 Top 50 Drugs, propublica.org") +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  geom_hline(yintercept = 0,
             size = 1,
             colour = "#333333") + coord_flip() +
  geom_label(
    aes(label = format(
      round(var, 0), big.mark = ",", scientific = FALSE
    )),
    hjust = 1,
    vjust = 0.5,
    colour = "white",
    fill = NA,
    label.size = NA,
    family = "Helvetica",
    size = 3
  ) +
  theme(
    panel.grid.major.x = element_line(color = "#cbcbcb"),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(1, 1, 1, 1, "cm"),
    legend.position = 'none',
    plot.title = element_text(size = 24),
    plot.subtitle = element_text(size = 16, face = "italic"),
    plot.caption = element_text(size = 6, hjust = 1)
  )

ggsave(
  'plots/topdrugs_avgpayment.png',
  plot = avgpayment_plot,
  units = "in",
  height = 7,
  width = 12,
  dpi = 300
)

```

### Plot: Average Total Physician Payment by Payment Type
```{r paytype_plot}

#Var =mean tot_payment per payment type
paytype_plot = drugs_interest %>% filter(received_payments == T) %>%
  select(drug_name,
         tot_meal_payments,
         tot_nonmeal_payments,
         tot_promo_payments) %>%
  rename(
    'Meals' = 'tot_meal_payments',
    'R&D/Consult/Travel' = 'tot_nonmeal_payments',
    'Promo. Lec.' = 'tot_promo_payments'
  ) %>%
  gather(key = 'payment_kind', value = 'amount', -drug_name) %>%
  group_by(payment_kind) %>%
  summarise(var = mean(amount)) %>%
  ggplot(data = ., aes(
    x = reorder(payment_kind, -var),
    y = var,
    fill = payment_kind
  )) +
  geom_bar(stat = "identity", position = 'dodge') +
  scale_fill_manual(values = c('#FAAB18' , '#1380A1', '#a13b0f')) +
  bbc_style() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  labs(title = "Type of Clinician Payments Across Top 20 Drugs",
       subtitle = "R&D/consult/travel and promotional lectures make up majority of payments",
       caption = "Data source: 2016 Top 50 Drugs, propublica.org") +
  geom_hline(yintercept = 0,
             size = 1,
             colour = "#333333") +
  geom_label(
    aes(label = format(
      round(var, 0), big.mark = ",", scientific = FALSE
    )),
    hjust = .5,
    vjust = 1,
    colour = "white",
    fill = NA,
    label.size = NA,
    family = "Helvetica",
    size = 5
  ) +
  theme(
    panel.grid.major.y = element_line(color = "#cbcbcb"),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(1, 1, 1, 1, "cm"),
    legend.position = 'none',
    plot.title = element_text(size = 24),
    plot.subtitle = element_text(size = 16, face = "italic"),
    plot.caption = element_text(size = 6, hjust = 1)
  ) + ylim(0, 300) +
  scale_colour_manual(
    labels = function(x)
      paste0(" ", x, "  ")
  )

ggsave(
  'plots/topdrugs_paymenttype.png',
  plot = paytype_plot,
  units = "in",
  height = 7,
  width = 12,
  dpi = 300
)

```

### Plot: Claims Per Payment
```{r claimspayment_plot}

#Organize Data
drug_data = drugs_interest %>% group_by(drug_name, payment_type) %>%
  select(drug_name, payment_type, n_claims_drug) %>%
  summarise(var = mean(n_claims_drug)) %>%
  spread(payment_type, var)

#Plot Data
claimspayment_plot = ggplot() +
  geom_segment(
    data = drug_data %>% mutate(
      start = NoPayments,
      end = ifelse(AllTypes > MealOnly, AllTypes, MealOnly)
    ),
    aes(
      x = start,
      xend = end,
      y = reorder(drug_name, (start - end)),
      yend = reorder(drug_name, (start - end))
    ),
    color = "#c9c4c2",
    size = 2
  ) +
  geom_point(
    data = drug_data %>% mutate(gap = ifelse(AllTypes > MealOnly, AllTypes, MealOnly) - NoPayments) %>%
      gather(payment_type, claims, AllTypes, MealOnly, NoPayments),
    aes(claims, reorder(drug_name, gap), color = payment_type),
    size = 4
  ) +
  scale_fill_manual(values = c('#FAAB18' , '#1380A1', '#a13b0f')) +
  bbc_style() +
  labs(title = "Average Clinician Rx Claim per Payment Type, By Drug",
       subtitle = "Notable differences between payment types",
       caption = "Data source: 2016 Top 50 Drugs, propublica.org") +
  theme(
    panel.grid.major.y = element_line(color = "#cbcbcb"),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(1, 1, 1, 1, "cm"),
    legend.position = 'right',
    plot.title = element_text(size = 24),
    plot.subtitle = element_text(size = 16, face = "italic"),
    plot.caption = element_text(size = 6, hjust = 1)
  )

ggsave(
  'plots/topdrugs_claimspayment.png',
  plot = claimspayment_plot,
  units = "in",
  height = 7,
  width = 12,
  dpi = 300
)


```

### Plot: Effect of Payment Amount
```{r paymentamt_group_plot}

paygroups_plot = drugs_interest %>% group_by(payment_amt_group) %>%
  summarise(var = mean(n_claims_drug)) %>%
  ggplot(data = ., aes(x=reorder(payment_amt_group,var), var, fill = payment_amt_group)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_manual(values = c('#dddddd', '#faab18', '#1380a1', '#601dfa', '#fa5b18')) +
  bbc_style() +
  labs(title = "Average Rx Claims Per Clinician, By Payment Group",
       subtitle = "Higher payments results in more prescriptions",
       caption = "Data source: 2016 Top 50 Drugs, propublica.org") +
  geom_hline(yintercept = 0,
             size = 1,
             colour = "#333333") +
  geom_label(
    aes(label = format(
      round(var, 0), big.mark = ",", scientific = FALSE
    )),
    hjust = .5,
    vjust = 1,
    colour = "white",
    fill = NA,
    label.size = NA,
    family = "Helvetica",
    size = 6
  ) +
  theme(
    panel.grid.major.y = element_line(color = "#cbcbcb"),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(1, 1, 1, 1, "cm"),
    legend.position = 'none',
        plot.title = element_text(size = 24),
    plot.subtitle = element_text(size = 16, face = "italic"),
    plot.caption = element_text(size = 6, hjust = 1)
  ) +
  scale_x_discrete(labels = c('$0', '$0< X <$100', '$100=< X <$500', '$500=< X <$1,000', 'X > $5,000'))

ggsave(
  'plots/topdrugs_paygroups.png',
  plot = paygroups_plot,
  units = "in",
  height = 7,
  width = 12,
  dpi = 300
)
  
```

### Plot: Effect of Payment Amount on OXYCONTIN
```{r claimspayment_oxycontin_plot}

oxycontin_paygroups_plot = drugs_interest %>% filter(drug_name == 'Oxycontin') %>% group_by(payment_amt_group) %>%
  summarise(var = mean(n_claims_drug)) %>%
  ggplot(data = ., aes(x=reorder(payment_amt_group,var), var, fill = payment_amt_group)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_manual(values = c('#dddddd', '#faab18', '#1380a1', '#601dfa', '#fa5b18')) +
  bbc_style() +
  labs(title = "Oxycontin: Average Rx Claims Per Clinician, By Payment Group",
       subtitle = "Higher payments results in more prescriptions",
       caption = "Data source: 2016 Top 50 Drugs, propublica.org") +
  geom_hline(yintercept = 0,
             size = 1,
             colour = "#333333") +
  geom_label(
    aes(label = format(
      round(var, 0), big.mark = ",", scientific = FALSE
    )),
    hjust = .5,
    vjust = 1,
    colour = "white",
    fill = NA,
    label.size = NA,
    family = "Helvetica",
    size = 6
  ) +
  theme(
    panel.grid.major.y = element_line(color = "#cbcbcb"),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(1, 1, 1, 1, "cm"),
    legend.position = 'none',
        plot.title = element_text(size = 24),
    plot.subtitle = element_text(size = 16, face = "italic"),
    plot.caption = element_text(size = 6, hjust = 1)
  ) +
  scale_x_discrete(labels = c('$0', '$0< X <$100', '$100=< X <$500', '$500=< X <$1,000', 'X > $5,000'))

ggsave(
  'plots/topdrugs_paygroups_OXYCONTIN.png',
  plot = oxycontin_paygroups_plot,
  units = "in",
  height = 7,
  width = 12,
  dpi = 300
)
  
```

### Plot: Payment Amount and Profit
```{r paymentamt_profit_plot}  

profits_plot = drugs_interest %>% mutate(payment_amt_group = factor(
  payment_amt_group,
  levels = c(
    "great_5k",
    "bet_500_1k",
    "bet_100_500",
    "bet_1_100",
    "nopayment"
  )
)) %>%
  filter(received_payments == T) %>%
  group_by(drug_name, payment_amt_group) %>%
  mutate(profit =  tot_claim_drug - tot_payments) %>%
  select(drug_name, payment_amt_group, profit) %>%
  summarise(var = sum(profit > 0) / n()) %>%
  ggplot(data = ., aes(payment_amt_group, y = var, fill = payment_amt_group)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  facet_wrap( ~ drug_name) +
  scale_fill_manual(values = c('#faab18', '#1380a1', '#601dfa', '#fa5b18')) +
  scale_x_discrete(
    labels = c(
      "bet_1_100" = "$0< X <$100",
      "bet_100_500" = '$100=< X <$500',
      "bet_500_1k" = '$500=< X <$1,000',
      'great_5k' = 'X > $5,000'
    )
  ) +
  bbc_style() +
  labs(title = "Clinician Payment Profitability Per Amount Group, By Drug",
       subtitle = "... like shooting fish in a barrel",
       caption = "Data source: 2016 Top 50 Drugs, propublica.org") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2)) +
  geom_hline(yintercept = 0,
             size = 1,
             colour = "#333333") + coord_flip() +
  geom_label(
    aes(label = scales::percent(round(var, 2))),
    hjust = 1,
    vjust = 0.5,
    colour = "white",
    fill = NA,
    label.size = NA,
    family = "Helvetica",
    size = 3
  ) +
  theme(
    text = element_text(size = 12),
    panel.grid.major.x = element_line(color = "#cbcbcb"),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(1, 1, 1, 1, "cm"),
    legend.position = 'none',
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
        plot.title = element_text(size = 24),
    plot.subtitle = element_text(size = 16, face = "italic"),
    plot.caption = element_text(size = 6, hjust = 1)
  )

ggsave(
  'plots/topdrugs_profits.png',
  plot = profits_plot,
  units = "in",
  height = 10,
  width = 14,
  dpi = 300
)

```